---
title: "HW3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load library & data
```{r}
rm(list=ls())
library(cluster)
library(fpc)
library(factoextra)
library(gridExtra)
library(conjoint)
# setwd("~/Desktop/AnalyticsDesign/project3")
setwd("~/Desktop/analytics design/HORSE")

load('GBA424 - Toy Horse Case Data.Rdata')

```
######## A. stimate the conjoint model at the individual level################
1.part-utilities

delete NA, profile 3,6,10,16
```{r}
df <- na.omit(conjointData)
```

create partworths
```{r}
partworths <- matrix(NA,nrow=nrow(respondentData),ncol=5)
for(i in 1:max(conjointData$ID)){ 
    partworths[i,]=lm(ratings~price+size+motion+style,data=df,subset=(ID==i))$coef
}
colnames(partworths) = c("Intercept","price","size","motion","style")
```


2.predictions for missing profiles
```{r}
fulldata <- conjointData
fullpartworths = matrix(rep(partworths,each=nrow(profilesData)),ncol=5)
intercept <- rep(1,nrow(conjointData))
x<-cbind(intercept,fulldata[,4:7])

fullratings = rowSums(x*fullpartworths) 

for (i in 1:nrow(fulldata)) {
    if (is.na(fulldata$ratings[i]) == T){
        fulldata$ratings[i] = fullratings[i]
    }
} 

fulldata
```


######### B.Conduct Benefit Segmentation via Cluster Analysis of Conjoint Part-Utilities##########

```{r}
clustTest = function(toClust,print=TRUE,scale=TRUE,maxClusts=15,seed=12345,nstart=20,iter.max=100){
    if(scale){ toClust = scale(toClust);}
    set.seed(seed);   # set random number seed before doing cluster analysis
    wss <- (nrow(toClust)-1)*sum(apply(toClust,2,var))
    for (i in 2:maxClusts) wss[i] <- sum(kmeans(toClust,centers=i,nstart=nstart,iter.max=iter.max)$withinss)
    ##gpw essentially does the following plot using wss above. 
    #plot(1:maxClusts, wss, type="b", xlab="Number of Clusters",ylab="Within groups sum of squares")
    gpw = fviz_nbclust(toClust,kmeans,method="wss",iter.max=iter.max,nstart=nstart,k.max=maxClusts) #alternative way to get wss elbow chart.
    pm1 = pamk(toClust,scaling=TRUE)
    ## pm1$nc indicates the optimal number of clusters based on 
    ## lowest average silhoutte score (a measure of quality of clustering)
    #alternative way that presents it visually as well.
    gps = fviz_nbclust(toClust,kmeans,method="silhouette",iter.max=iter.max,nstart=nstart,k.max=maxClusts) 
    if(print){
        grid.arrange(gpw,gps, nrow = 1)
    }
    list(wss=wss,pm1=pm1$nc,gpw=gpw,gps=gps)
}
##Runs a set of clusters as kmeans
##Arguments:
##  toClust, data.frame with data to cluster
##  nClusts, vector of number of clusters, each run as separate kmeans 
##  ... some additional arguments to be passed to clusters
##Return:
##  list of 
##    kms, kmeans cluster output with length of nClusts
##    ps, list of plots of the clusters against first 2 principle components
runClusts = function(toClust,nClusts,print=TRUE,maxClusts=15,seed=12345,nstart=20,iter.max=100){
    if(length(nClusts)>4){
        warning("Using only first 4 elements of nClusts.")
    }
    kms=list(); ps=list();
    for(i in 1:4){
        kms[[i]] = kmeans(toClust,nClusts[i],iter.max = iter.max, nstart=nstart)
        ps[[i]] = fviz_cluster(kms[[i]], geom = "point", data = toClust) + ggtitle(paste("k =",nClusts[i]))
        
    }
    library(gridExtra)
    if(print){
        tmp = marrangeGrob(ps, nrow = 2,ncol=2)
        print(tmp)
    }
    list(kms=kms,ps=ps)
}

##Plots a kmeans cluster as three plot report
##  pie chart with membership percentages
##  ellipse plot that indicates cluster definitions against principle components
##  barplot of the cluster means
plotClust = function(km,toClust,discPlot=FALSE){
    nc = length(km$size)
    if(discPlot){par(mfrow=c(2,2))}
    else {par(mfrow=c(3,1))}
    percsize = paste(1:nc," = ",format(km$size/sum(km$size)*100,digits=2),"%",sep="")
    pie(km$size,labels=percsize,col=1:nc)
    
    clusplot(toClust, km$cluster, color=TRUE, shade=TRUE,
             labels=2, lines=0,col.clus=1:nc); #plot clusters against principal components
    
    if(discPlot){
        plotcluster(toClust, km$cluster,col=km$cluster); #plot against discriminant functions ()
    }
    rng = range(km$centers)
    dist = rng[2]-rng[1]
    locs = km$centers+.05*dist*ifelse(km$centers>0,1,-1)
    bm = barplot(km$centers,beside=TRUE,col=1:nc,main="Cluster Means",ylim=rng+dist*c(-.1,.1))
    text(bm,locs,formatC(km$centers,format="f",digits=1))
}

checks = clustTest(partworths,print=TRUE,scale=TRUE,maxClusts = 15,seed=12345,nstart=20,iter.max=100)
clusts=runClusts(partworths,c(2,3,4,5),print=TRUE,maxClusts=15,seed=12345,nstart=20,iter.max=100)
for(i in 1:4) plotClust(clusts[[1]][[i]],partworths)
```
optimal number of cluster=3

############ C.	Conduct a priori segmentation ##############

by age
```{r}
fulldf <- merge(df,respondentData,by='ID')
summary(lm(ratings~(price+size+motion+style)*age,data=fulldf))
summary(lm(ratings~price+size+motion+style,data=fulldf,subset=(age==0))) # 2 yrs old
summary(lm(ratings~price+size+motion+style,data=fulldf,subset=(age==1))) # 3-4 yrs old

```


by gender
```{r}
summary(lm(ratings~(price+size+motion+style)*gender,data=fulldf))
summary(lm(ratings~price+size+motion+style,data=fulldf,subset=(gender==0))) # male 
summary(lm(ratings~price+size+motion+style,data=fulldf,subset=(gender==1))) #female  

```


#################D. Simulate market shares for different product-line scenarios##############
```{r}
rankdata = as.data.frame(matrix(nrow = 200, ncol = 16))
colnames(rankdata) = paste0('profile',1:16)
for(j in 1:200){
    rankdata[j,]=fulldata$ratings[fulldata$ID==j]
}
#simulates decisions for a market scenario using first choice disaggregate choice model
##Arguments:
##  scen, the scenario consisting of a set of options (columns) to simulate decisions over. Takes variable names, numerical indexes, or boolean
##  data, matrix or data.frame of rank order data (1 best) with nrow=#cases and ncol=#options
##  ascend=TRUE, whether the rank order is ascending meaning lower ranks ar more preferred. 
##Return:
##  data.frame of decisions with nrow=nrow(data), ncol=length(scen) 
##    containing 1 or 0 and each row summing to 1 (one-hot encoding or dummy coded)
simFCDecisions = function(scen,data,ascend=FALSE){ 
  inmkt = data[,scen] #construct the subsetted matrix of options
  bestOpts <- matrix(NA,nrow=200,ncol=ncol(inmkt))
  if(ascend){ #if ranks 1 is best 
    for(i in 1:200){
      bestOpts[i,]=as.integer(inmkt[i,]==min(inmkt[i,]))
      numoftie=sum(bestOpts[i,])
      bestOpts[i,]=bestOpts[i,]/numoftie
    }
   } else { #else the best rank is the largest number
    for(i in 1:200){
      bestOpts[i,]=as.integer(inmkt[i,]==max(inmkt[i,]))
      numoftie=sum(bestOpts[i,])
      bestOpts[i,]=bestOpts[i,]/numoftie
    }
   }
  bestOpts=as.data.frame(bestOpts)
  #ret = as.data.frame(model.matrix(~0+as.factor(bestOpts)))#fill to set of options marked 0 or 1
  names(bestOpts) = names(inmkt)
  bestOpts
}
```


```{r}
#calculates shares given decisions matrix
##Argument:
##  decisions, matrix or data.frame containing decisions with nrow=#cases and ncol=#options
##Return:
##  a vector of shares
#note this formulation can handle 1/0 decisions, split decisions, and quantity decisions
calcUnitShares = function(decisions){
  colSums(decisions)/sum(decisions) #assumes that total decisions is market size
}

#simulates decisions for a market scenario using first choice disaggregate choice model
##Arguments:
##  data, matrix or data.frame of rank order data (1 best) with nrow=#cases and ncol=#options
##  scen, the scenario consisting of a set of options (columns) to simulate decisions over. Takes variable names, numerical indexes, or boolean
##  ascend=TRUE, whether the rank order is ascending meaning lower ranks ar more preferred. 
##Return:
##  a vector of shares
# note this formulation can handle 1/0 decisions, split decisions, and quantity decisions
simFCShares=function(scen,rankdata,ascend=F){
  decs = simFCDecisions(scen,rankdata,ascend) #determine decisions
  calcUnitShares(decs) #calculate shares and return
}


#i.e
statusQuo=c(5,7,13)
out = simFCDecisions(statusQuo,rankdata)
simFCShares(statusQuo,rankdata) # 0.22*4000*(111.99-33)+0.21*4000*(111.99-33)-20000*2 #0.57*4000*(111.99-41)-20000

```


```{r}
##Set up scenarios.  Here in this case we assume that companies just have one time to take action, no second,third,.... chance to react(because if react continuously, we have to calculate new cost for new products lauching, here we have 20000/3 for products not in existing set, but we do not have cost data for new "not existing products").
scens = list()
scens[[1]]=c(7,3,13,15)      
scens[[2]]=c(7,3,15)   
scens[[3]]=c(7,3,13)        
scens[[4]]=c(7,13,15) 

### they betray us first, they only have one product to choose to drop price, we react with differnt low-price products combination, so we list all possibilitis except all low prices###
scens[[5]]=c(8,3,13,15)      
scens[[6]]=c(8,3,15)   
scens[[7]]=c(8,3,13)        
scens[[8]]=c(8,13,15) 
scens[[9]]=c(8,4,13,15)
scens[[10]]=c(8,3,14,15)
scens[[11]]=c(8,3,13,16)
scens[[12]]=c(8,3,14,16)
scens[[13]]=c(8,4,13,16)
scens[[14]]=c(8,4,14,15)
scens[[15]]=c(8,4,15)
scens[[16]]=c(8,3,16)
scens[[17]]=c(8,4,13)
scens[[18]]=c(8,3,14)
scens[[19]]=c(8,13,16)
scens[[20]]=c(8,14,15)
## we betray them first, we could drop prices of only one, or 2, or all 3 products. so we list all possibilitis ##
scens[[21]] = c(7,4,15)
scens[[22]] = c(7,3,16)
scens[[23]] = c(7,4,16)  
scens[[24]] = c(7,4,13)
scens[[25]] = c(7,3,14) 
scens[[26]] = c(7,4,14) 
scens[[27]] = c(7,14,15) 
scens[[28]] = c(7,13,16)
scens[[29]] = c(7,14,16)
scens[[30]] = c(7,4,13,15)
scens[[31]] = c(7,3,14,15)
scens[[32]] = c(7,3,13,16)
scens[[33]] = c(7,4,14,15)
scens[[34]] = c(7,4,13,16) 
scens[[35]] = c(7,3,14,16)
scens[[36]] = c(7,4,14,16) 


# Drop all products prices together. 
scens[[37]] = c(8,4,14,16)
scens[[38]] = c(8,4,14)
scens[[39]] = c(8,14,16)
scens[[40]] = c(8,4,16)

share <- sapply(scens,simFCShares,rankdata)

profitMargin = c(0,0,111.99-29,95.99-29,111.99-33,95.99-33,111.99-41,95.99-41,0,0,0,0,111.99-33,95.99-33,111.99-41,95.99-41)
profit <- c()
for (j in 1:length(scens)){
  if (13 %in% scens[[j]] | 14 %in% scens[[j]]){
    productProfit <- -(length(scens[[j]])-1)*20000 - (length(scens[[j]])-1-1)*20000/3 #fixed cost
  }
  else{
    productProfit <- -(length(scens[[j]])-1)*20000 - (length(scens[[j]])-1)*20000/3 #fixed cost
  }
  for (i in 2:length(scens[[j]])){
     productProfit <- productProfit + share[[j]][[i]]*4000*profitMargin[scens[[j]][[i]]]
    }
  profit <- append(profit, productProfit)
}
profit

profitCompetSet <- c()
for (j in 1:length(scens)){
  profitCompet <- share[[j]][[1]]*4000*profitMargin[scens[[j]][[1]]] - 20000 # assume compet have same cost strycture as us
  profitCompetSet <- append(profitCompetSet, profitCompet)
}
profitCompetSet
```











#############appendix - write output into csv###############
```{r}
outputBothHighPr = as.data.frame(matrix(nrow = 4, ncol = 4))
scs <- c(7,3,13,15)
colnames(outputBothHighPr) = paste0('profile',c(7,3,13,15))
outputBothHighPr=cbind('scenario'=1:4,outputBothHighPr)
outputBothHighPr=cbind(outputBothHighPr,profit=profit[1:4])
outputBothHighPr=cbind(outputBothHighPr,profitCompet=profitCompetSet[1:4])
for (j in 1:length(scs)){
  sharePerProf <- c()
  for (i in 1:4){
    if (scs[j] %in% scens[[i]]){
      sharePerProf <- append(sharePerProf, share[[i]][[which(scens[[i]] == scs[j])]])
    }
    else {
      sharePerProf <- append(sharePerProf, NA)
    }
  }
  outputBothHighPr[1+j] <- sharePerProf
}

write.csv(outputBothHighPr,"outputBothHighPr.csv", row.names = FALSE)
read.csv("outputBothHighPr.csv")

outputTheyBetray = as.data.frame(matrix(nrow = 16, ncol = 7))
scs <- c(8,3,4,13,14,15,16)
colnames(outputTheyBetray) = paste0('profile',scs)
outputTheyBetray=cbind('scenario'=5:20,outputTheyBetray)
outputTheyBetray=cbind(outputTheyBetray,profit=profit[5:20])
outputTheyBetray=cbind(outputTheyBetray,profitCompet=profitCompetSet[5:20])
for (j in 1:length(scs)){
  sharePerProf <- c()
  for (i in 5:20){
    if (scs[j] %in% scens[[i]]){
      sharePerProf <- append(sharePerProf, share[[i]][[which(scens[[i]] == scs[j])]])
    }
    else {
      sharePerProf <- append(sharePerProf, NA)
    }
  }
  outputTheyBetray[1+j] <- sharePerProf
}

write.csv(outputTheyBetray,"outputTheyBetray.csv", row.names = FALSE)
read.csv("outputTheyBetray.csv")


outputWeBetray = as.data.frame(matrix(nrow = 16, ncol = 7))
scs <- c(7,3,4,13,14,15,16)
colnames(outputWeBetray) = paste0('profile',scs)
outputWeBetray=cbind('scenario'=21:36,outputWeBetray)
outputWeBetray=cbind(outputWeBetray,profit=profit[21:36])
outputWeBetray=cbind(outputWeBetray,profitCompet=profitCompetSet[21:36])
for (j in 1:length(scs)){
  sharePerProf <- c()
  for (i in 21:36){
    if (scs[j] %in% scens[[i]]){
      sharePerProf <- append(sharePerProf, share[[i]][[which(scens[[i]] == scs[j])]])
    }
    else {
      sharePerProf <- append(sharePerProf, NA)
    }
  }
  outputWeBetray[1+j] <- sharePerProf
}

write.csv(outputWeBetray,"outputWeBetray.csv", row.names = FALSE)
read.csv("outputWeBetray.csv")

outputBothLowPr = as.data.frame(matrix(nrow = 4, ncol = 4))
scs <- c(8,4,14,16)
colnames(outputBothLowPr) = paste0('profile',scs)
outputBothLowPr=cbind('scenario'=37:40,outputBothLowPr)
outputBothLowPr=cbind(outputBothLowPr,profit=profit[37:40])
outputBothLowPr=cbind(outputBothLowPr,profitCompet=profitCompetSet[37:40])
for (j in 1:length(scs)){
  sharePerProf <- c()
  for (i in 37:40){
    if (scs[j] %in% scens[[i]]){
      sharePerProf <- append(sharePerProf, share[[i]][[which(scens[[i]] == scs[j])]])
    }
    else {
      sharePerProf <- append(sharePerProf, NA)
    }
  }
  outputBothLowPr[1+j] <- sharePerProf
}

write.csv(outputBothLowPr,"outputBothLowPr.csv", row.names = FALSE)
read.csv("outputBothLowPr.csv")
```


write into csv.

